---
title: "Analysing local authority carbon emissions"
output: html_notebook
---

# Task

Understanding how carbon emissions in local authorities has changed since 2005.

# Data sources

https://www.gov.uk/government/statistics/uk-local-authority-and-regional-carbon-dioxide-emissions-national-statistics-2005-to-2019

# Data preparation and processing

# Analysis

```{r}
library(tidyverse)  #helps wrangle data
library(ggplot2)  #helps visualize data
```

Load data:

```{r}
rawcarbon <- read_csv("csv/2005-19_Local_Authority_CO2_emissions.csv")
```

```{r}
# Let's start by picking out a single region, York
singleregion <- filter(rawcarbon, `Local Authority Code` == 'E06000014')
# Then let's check what years are in our dataset
unique(singleregion$`Calendar Year`)
# Visual check that summary total matches original dataset
test <- singleregion %>% 
  filter(`Calendar Year` == '2005') 
sum(test$`Territorial emissions (kt CO2)`) 
```

# Let's summarise by total emmission per year across all categories for one LA
```{r}
singleregionsummary <- singleregion %>%
  group_by(`Calendar Year`) %>%
  summarise(Total = sum(`Territorial emissions (kt CO2)`, na.rm = TRUE))
ggplot(singleregionsummary, aes(x = `Calendar Year`, y = Total)) + 
  geom_line(aes(), size = 1)  + 
  theme_minimal() + 
  xlab("Year") + ylab("Territorial emissions (kt CO2)")
```

# TODO: next I need to do the line graph for a single region but show the change per territorial emissions group

# TODO - all councils summary

```{r}
tmpcarbon <- rawcarbon %>% 
  group_by(`Local Authority Code`, `Calendar Year`) %>% 
  summarise(Total = sum(`Territorial emissions (kt CO2)`, na.rm = TRUE))
```
```{r}

library("reshape2")
carbonWide <- dcast(tmpcarbon, `Local Authority Code` ~ `Calendar Year`, value.var="Total")
```

# TODO 

 * Validate sum in carbonWide
 * Include Local Authority Name in wide data for labelling
 * Plot trend graph for all
